---
title: "Reproducible Research: Peer Assessment 1"
author: "Simone Sinkovec"
date: "4/12/2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Load dataset 'Activity' in appropriate dataframe

```{r}
activityMonitoring <- read.csv("activity.csv", header = TRUE, sep = ",", dec = ".")
head(activityMonitoring)
str(activityMonitoring)
names(activityMonitoring)
summary(activityMonitoring)
```

## What is mean total number of steps taken per day?

Assignment part 1: What is mean total number of steps taken per day?
Subset & calculate sum per day, put it in histogram
```{r}
daily_steps <- aggregate(steps ~ date, data = activityMonitoring, FUN = sum)
hist(daily_steps$steps, main = "Total number of steps taken per day", xlab = "Steps per Day")
```

Calculate mean and median of steps per day
```{r}
meanSteps <- mean(daily_steps$steps)
mean(daily_steps$steps)

medianSteps <- median(daily_steps$steps)
median(daily_steps$steps)
```

## What is the average daily activity pattern?

Assignment part 2: What is the average daily activity pattern?
```{r}
library(ggplot2)
stepsInterval <- aggregate(steps ~ interval, data = activityMonitoring, FUN = mean)
  g <- ggplot(stepsInterval, aes(interval, steps))
  g + geom_line()
    stepsInterval[which.max(stepsInterval$steps), ]
```

## Imputing missing values

Assignment part 3: Imputing missing values
Calculate total number of missing rows
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
sum(!complete.cases(activityMonitoring))
activityMonitoring$steps[is.na(activityMonitoring$steps)] = mean(activityMonitoring$steps, na.rm=TRUE)
daily_steps2 <- aggregate(steps ~ date, data = activityMonitoring, FUN = sum)
par(mfrow = c(1,2))
hist(daily_steps$steps, main = "Total number of steps taken per day", xlab = "Steps per Day")
hist(daily_steps2$steps, main = "Total number of steps taken per day (NA = mean)", xlab = "Steps per Day")
```

Calculate mean and median of steps per day
```{r}
meanSteps2 <- mean(daily_steps2$steps)
medianSteps2 <- median(daily_steps2$steps)
library(xtable)
result <- data.frame(c(meanSteps, medianSteps), c(meanSteps2, medianSteps2))
```
The replacement of missing values with the mean, doesn't have much effect on histogram, mean and median.

## Are there differences in activity patterns between weekdays and weekends?

Assignment part 4:
Plot 'Differences in activity patterns between weekdays and weekends'
```{r}
activityMonitoring$date <- as.Date(activityMonitoring$date, "%Y-%m-%d")
library(dplyr)
activityMonitoring$day <- weekdays(activityMonitoring$date)
activityMonitoring <- mutate(activityMonitoring, weekSeparation = ifelse((day == "Sunday") | (day == "Saturday"), "weekend", "weekday"))
activityMonitoring$weekSeparation <- as.factor(activityMonitoring$weekSeparation)
stepsIntervalWeeksepation <- aggregate(steps ~ interval + weekSeparation, data = activityMonitoring, mean)
g <- ggplot(stepsIntervalWeeksepation, aes(interval, steps, color = weekSeparation))
g + geom_line() + facet_grid(weekSeparation~.) + ggtitle("Differences in activity patterns between weekdays and weekends") + xlab("5 Minute Interval") + ylab("Average amount of steps")